---
title: "Network analysis tutorial"
author: "Jiaxin Deng"
date: "2019/5/20"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Brief introduction to network analysis


A network is a set of nodes connected by a set of edges.

Several packages are used in the network analysis, including `network`, `statnet`, `igraph` and `qgraph`.

`qgraph` was developed in the context of psychometrics approach by Dr. Sacha Epskamp and colleagues in 2012. For more details, please click this following link for the paper published in *Journal of Statistical Softare*:

<https://www.jstatsoft.org/article/view/v048i04>

This package can create graphs to visualize the statistics in different layout modes based on different correlation matrices, such as polychoric correlation, partial correlation.


# Example code

Here is some example code to show how to conduct a network analysis.


## Estimating networks 

### How to make a graph

Here is the following steps to make a gragh using `qgraph`.

Take `big5` data as an example. This is a dataset of the Big five personality traits assessed on 500 psychology students.

Firstly, `qgraph` package should be activated using `library()`

```{r library()}
library(qgraph)
```

And then, data need to be imported in the current R project.

```{r get big5 data}
data(big5)
```


